---
title: Simulating Dynamics on Networks
author: Neil MacLaren
date: "`r format(Sys.Date(), '%b %d, %Y')`"
output:
  pdf_document:
    latex_engine: pdflatex
---

```{r echo = FALSE, eval = TRUE}
.libPaths("/home/neil/R/library")
## lout <- 100
```

*sdn* is a package for simulating dynamics on networks. The package grew out of a need to do the same tasks with both deterministic and stochastic simulations of the same model, but has expanded as I have come to rely on it more. The current goal is to support numerical investigation of $N$ dynamical equations coupled via a complex network. As such, it supports ordinary and stochastic differential equations, different coupling functions, and phase space investigation.

In this document I intend to discuss a variety of topics related to dynamics on networks and how I address them using this package and the more fundamental packages upon which I most rely: *deSolve* and *igraph*. 

# Fundamentals

One way to look at dynamics on networks is to imagine that there is some dynamical (i.e., changing in time) process occuring on the nodes of a network, and that each dynamical process influences each other dynamical process in some way through the network edges. We can write this idea generically as
\begin{equation}
  \frac{dx_i}{dt} = F(x_i) + G(x_i, x_j),
\end{equation}
where $F(x_i)$ is the dynamics happening on each node and $G(x_i, x_j)$ is a coupling function. The function $F(x_i)$ defines what the node does in isolation, whereas the function $G(x_i, x_j)$ defines how two nodes influence each other. Thus, the change in state of the $i$th node is a function of some (often nonlinear) internal process and the state of its network neighbors. 

Consider as an example the following dynamical model
\begin{equation}
  \frac{dx_i}{dt} = -(x_i - r_1)(x_i - r_2)(x_i - r_3) + D\sum_{j=1}^N A_{i,j}x_j.
\end{equation}
Here, $F(x_i) = -(x_i - r_1)(x_i - r_2)(x_i - r_3)$, where $x_i$ is a numeric state of the $i$th node and $r_1 < r_2 < r_3$ are parameters which control the position of the equilibria in the absence of coupling. The coupling function $G(x_i, x_j) = D\sum_{j=1}^N A_{i,j}x_j$---we'll come back to this below. Before we do, let's get a sense of what's happening with $F(x_i)$.

The node dynamics $F(x_i) = -(x_i - r_1)(x_i - r_2)(x_i - r_3)$ is known as a double-well model. If we plot several possible initial values of $x$ against $t$ we get something like this:

```{r echo = FALSE, eval = TRUE, fig.width = 4, fig.height = 4}
library(deSolve)
library(sfsmisc)

.dw <- list(
    r = c(1, 3, 5)
)
dw <- function(t, x, params) {
    with(params, {
        dx <- -(x - r[1])*(x - r[2])*(x - r[3])
        return(list(c(dx)))
    })
}

times <- seq(0, 1, length.out = 500)
X <- ode(2, times, dw, .dw)
xinits <- seq(5)
n <- 8
Xs <- lapply(xinits, ode, times = times, func = dw, parms = .dw)

palette("Dark 2")
par(mar = c(5, 5, 0.5, 0.5))
plot(
    NULL, xlim = range(times), ylim = range(.dw$r),
    xlab = "t", ylab = "x", font.lab = 3, cex.lab = 2, cex.lab = 1.75,
    axes = FALSE
)
eaxis(1)
eaxis(2)
invisible(mapply(function(X, i) lines(X, col = i, lwd = 2), Xs, xinits))
```
